### 2023-04-05 submission/manuscript.Rmd
Through uneven sample size analysis

Each simulated condition was replicated 5 times in WNWN and the paper reports the mean and standard deviation of the replicate clustering accuracies. The relatively small number of replicates accounts for the jerkiness of the lines in the original Figures 4 and 5 (e.g. the Bray-Curtis distances calculated on the DEeqVS transformed data). A better approach would have been to use 100 replicates as this would reduce the dependency of the results on the random number generator's seed. Furthermore, because the accuracies are unlikely to be symmetric around a mean at larger accuracy values the median and 95% confidence intervals or intraquartile range should have been reported. To test the effect of increasing the number of replicates, I pulled apart the code in `simulation-cluster-accuracy/simulation-cluster-accuracy-server.Rmd` into individual R and bash scripts that were executed using a Snakemake workflow with the same Conda environment as above. This was necessary since the number of simulated conditions because of this change increased 20-fold. Such intense data processing was not practical within a single R-flavored markdown document. Again, the observed results were qualitiatively similar to those generated using the single R-flavored markdown file (Figures S4 (**pam_subsample15_fig_4.pdf**) and S5(**pam_subsample_fig_5.pdf**) ). The increased number of replications resulted in smoother lines and allowed me to present empirical 95% confidence intervals. For all analyses in the remainder of this paper, I used 100 randomized replicates per condition.

@@

In previous analyses I have observed that not using rarefaction can lead to falsely detecing differences between communities when sampling effort is confounded with the treatment group [XXXXXXX]. Such situations have been observed when comparing communities at different body parts where one site is more likely to generate contaminating sequence reads from the host [XXXXXXX]. My previous analyses showed that rarefaction did the best job of controlling the rates of false detection (i.e., Type I errors) and false rejection (i.e., Type II errors).

To determine whether this result was replicated with the WNWN simulation framework, I created a skewed sampling distribution using both the GlobalPatterns and Log-distributed sequence distributions. To skew the sample counts the sequencing depth of samples from one treatment group were drawn from below the median number of sequences of the sampling distribution and those for the second treatment group were from above the median. To assess Type I errors, I compared the clustering accuracies using an effect size of 1.0 using both the skewed and unskewed sampling distributions (Figure **cluster_skew_compare_i.pdf**). For the case when the effect size was 1.0, the samples should have only been assigned to one cluster. However, each of the clustering methods forced the samples into two clusters. So, when there are two groups of 40 samples that do not differ, the best a method could do would be to correctly assign 41 of the 80 samples for an accuracy of `r 41/80`. As has been seen previously, the Type I error does not vary by method when the sequencing depth is not skewed. Yet, when the sequencing depth was skewed, rarefaction was the most consistent normalization method for controlling Type I errors.

@@

The analysis of the GlobalPatterns-based simulations in WNWN used clustering accuracy to assess the effect of different normalization procedures, distance calculations, and clustering methods. There has been controversy over the meaning of clustering samples (i.e., enterotypes) including whether such clustering should be done on ecological distances or sequence counts and the biological interpretation of such clusters [XXXXXXXXX]. One notable challenge with using clustering accuracy as the dependent variable is that the clustering methods force the samples into one of two clusters. For the case where the effect size was 1.0, it was impossible for all 80 samples to be assigned to a single cluster. As has already been demonstrated, an additional challenge with using clustering accuracy is that there is sensitivity to the clustering method that a researcher selects. Given the apparent interaction between the various variables and clustering method with clustering accuracy, an alternative approach is needed. In fact, a more common analysis of distance matrices would be to use a non-parametric analysis of variance test of the various distance matrices. Therefore, I subjected each of the distance matrices to such a test using `adonis2`, a function from the vegan R package that implements this test. This analysis allowed me to quantify the probability of falsely rejecting a null hypothesis (i.e., Type I error) when the effect size was 1.0 and the power to detect a difference when the effect sizes were larger than 1.0. Such an analysis was not possible in the original study since only 5 replicates of each simulation were performed.

>>>>

Each simulated condition was replicated 5 times in WNWN and the paper reports the mean and standard deviation of the replicate clustering accuracies. The relatively small number of replicates accounts for the jerkiness of the lines in the original Figures 4 and 5 (e.g. the Bray-Curtis distances calculated on the DEeqVS transformed data). A better approach would have been to use 100 replicates as this would reduce the dependency of the results on the random number generator's seed. By increasing the number of replicates it was also possible to compare the probability of falsely and correctly clustering samples from the same and different treatment groups together (see points 9-11, below). Because the accuracies are unlikely to be symmetric around a mean at larger accuracy values the median and 95% confidence intervals or intraquartile range should have been reported. To test the effect of increasing the number of replicates, I pulled apart the code in `simulation-cluster-accuracy/simulation-cluster-accuracy-server.Rmd` into individual R and bash scripts that were executed using a Snakemake workflow with the same Conda environment as above. This was necessary since the number of simulated conditions because of this change increased 20-fold. Such intense data processing was not practical within a single R-flavored markdown document. Again, the observed results were qualitiatively similar to those generated using the single R-flavored markdown file (Figures S4 (**pam_subsample15_fig_4.pdf**) and S5(**pam_subsample_fig_5.pdf**) ). The increased number of replications resulted in smoother lines and allowed me to present empirical 95% confidence intervals. For all analyses in the remainder of this paper, I used 100 randomized replicates per condition.

@@

In previous analyses I have observed that not using rarefaction can lead to falsely detecing differences between communities when sampling effort is confounded with the treatment group [XXXXXXX]. Such situations have been observed when comparing communities at different body parts where one site is more likely to generate contaminating sequence reads from the host [XXXXXXX]. My previous analyses showed that rarefaction did the best job of controlling the rates of false detection (i.e., Type I errors) and maintaining the statistical power to detect differences (i.e., 1-rate of Type II errors) of differences between groups of samples.

To determine whether this result was replicated with the WNWN simulation framework, I created a skewed sampling distribution using both the GlobalPatterns and Log-distributed sequence distributions. To skew the sample counts the sequencing depth of samples from one treatment group were drawn from below the median number of sequences of the sampling distribution and those for the second treatment group were from above the median. To assess Type I errors, I compared the clustering accuracies using an effect size of 1.0 using both the skewed and unskewed sampling distributions (Figure **cluster_skew_compare_i.pdf**). For the case when the effect size was 1.0, the samples should have only been assigned to one cluster. However, each of the clustering methods forced the samples into two clusters. So, when there are two groups of 40 samples that do not differ, the best a method could do would be to correctly assign 41 of the 80 samples for an accuracy of `r 41/80`. The Type I error did not vary by method when the sequencing depth was not skewed. Yet, when the sequencing depth was skewed, rarefaction was the most consistent normalization method for controlling Type I errors. At larger effect sizes the power to detect differences increased when the sequencing depth was skewed (Figure **cluster_skew_compare_ii.pdf**). At the small effect size of 1.15, with the exception of the Poisson distance data, which generally clustered performed poorly, the rarefied data generated the highest accuracy clusters regardless of whether the data were skewed. Although the skewed simulation is extreme, it highlights the ability of rarefaction to control Type I errors while maintaining high power.

@@

The analysis of the GlobalPatterns-based simulations in WNWN used clustering accuracy to assess the effect of different normalization procedures, distance calculations, and clustering methods. There has been controversy over the meaning of clustering samples (i.e., enterotypes) including whether such clustering should be done on ecological distances or sequence counts and the biological interpretation of such clusters [XXXXXXXXX]. As described in the previous point, one notable challenge with using clustering accuracy as the dependent variable is that the clustering methods force the samples into one of two clusters. For the case where the effect size was 1.0, it was impossible for all 80 samples to be assigned to a single cluster. As has already been demonstrated in point 5, an additional problem wiht clustering is the sensitivity to the chosen method. A more common analysis of distance matrices is to use a non-parametric analysis of variance test of the various distance matrices (i.e., AMOVA, PERMANOVA, adonis)[XXXXXXXXXXX]. Therefore, I subjected each of the distance matrices to such a test using `adonis2`, a function from the vegan R package that implements this test.
